---
title: "Financial_Engineering_HA3"
author: "Burzler R., Krejci T., Wittmann M., Zaborskikh E."
date: "April 7, 2019"
header-includes:
   - \usepackage{tikz,pgfplots}
   - \usepackage{amsmath}
   - \usepackage{amssymb}
   - \usepackage{epstopdf}
   - \usepackage{mathtools}
   - \usepackage{calculator}
   - \usepackage{a4wide}
   - \pgfplotsset{compat=1.12}
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(pacman)
```

```{r packages, echo=FALSE}
p_load("tidyr","ggplot2","tictoc")
```

## Exercise 1

Suppose we have short rate $r_t$ following a Vasicek process with $r_0 = 0.02$, $\bar{r} = 0.05$, $\gamma = 0.2$, $\sigma = 0.015$. Take $T = 1$ and number of steps equal to $250$.

We generate 40 paths and plot our results (you can use exact simulation or Euler discretization). Now keeping everything fixed we change the parameter $\gamma$ to 5. Finally, we visualize the generated paths in a separate plots:
```{r Vasicek}
# Generation of Path following the Vasicek Model
paths.VAS<-function(r0=.02, r=.05, gamma=.2, sigma=.015, steps=250, n.paths=40){
  Tcap <- 1
  h <- Tcap/steps
  r_t <- as.data.frame(matrix(rep(c(r0,rep(NA,steps-1)),n.paths), ncol = n.paths))
  names(r_t) <- paste0(rep('Path.',n.paths), c(1:n.paths))    
    for(j in 1:n.paths) {
      for(i in 2:250) {
        r_t[i,j] <- r_t[i-1,j]*exp(-gamma*h)+r*(1-exp(-gamma*h)) +
                    sigma*sqrt((1-exp(-2*gamma*h))/(2*gamma))*rnorm(1,0,1)
        }
    }
  A <- cbind(c(1:steps), r_t)
  names(A)[1] <- 'Step'
  return(A)
}

## Evaluation
set.seed(2019)
pl.dt.vas <- gather(paths.VAS(), 'Path', 'Value', 2:41)
set.seed(2019)
pl.dt.vas5 <- gather(paths.VAS(gamma=5), 'Path', 'Value', 2:41)

ggplot(pl.dt.vas,aes(Step,Value,group=Path))+
  geom_line(aes(colour=Path))+
  theme(legend.title = element_blank())+
  theme(legend.position = "none")+
  ggtitle('Vasicek paths - gamma=0.2')
```
In case of a very low speed of mean-reversion ($\gamma =0.2$), we visually see that the process does not come back from the initial $r_0 = 0.02$ to the long-run mean $\bar{r} =0.05$ within the 250 observations generated. Note that we can also observe some negative values for $r_t$ in this case.


\newpage

In case we change the speed of mean-reversion to be quite high ($\gamma =5$), we visually see that the process comes back quite fast from the initial $r_0 = 0.02$ to the long-run mean $\bar{r} =0.05$.

```{r plot_Vasicek5}
ggplot(pl.dt.vas5,aes(Step,Value,group=Path))+
  geom_line(aes(colour=Path))+
  theme(legend.title = element_blank())+
  theme(legend.position = "none")+
  ggtitle('Vasicek paths - gamma=5')
```

\newpage
## Exercise 2

Suppose we have a short rate $r_t$ following CIR process with $r_0 = 0.02$, $\bar{r} = 0.05$, $\gamma = 1.2$, $\alpha = 0.04$. Take $T = 1$ and number of steps equal to 250. We use Euler discretization using truncation or reflection to avoid negative values. Finally, we generate 40 paths and plot our results:
```{r CIR}
## Path Generation
paths.CIR<-function(r0=.02,r=.05, gamma=.2,alpha=.04,steps=250,n.paths=40,negs='truncate'){
  Tcap <- 1
  h <- Tcap/steps
  r_t <- as.data.frame(matrix(rep(c(r0,rep(NA,steps-1)),n.paths), ncol = n.paths))
  names(r_t) <- paste0(rep('Path.',n.paths), c(1:n.paths))    
  if(negs=='truncate') {
      for(j in 1:n.paths) {
        for(i in 2:250){
          r_t[i,j] <- gamma*r*h+(1-gamma*h)*max(r_t[i-1,j],0) + 
                      sqrt(alpha*max(r_t[i-1,j],0)*h)*rnorm(1,0,1)
          }
      }
  } else if(negs=='reflect'){
      for(j in 1:n.paths){
        for(i in 2:250){
          r_t[i,j] <- gamma*r*h+(1-gamma*h)*max(r_t[i-1,j],-r_t[i-1,j]) +
                      sqrt(alpha*max(r_t[i-1,j],-r_t[i-1,j])*h)*rnorm(1,0,1)
          }
      }
  } else {stop("Options for handling negative r's are 'truncate' and 'reflect'")}
      
  A <- cbind(c(1:steps), r_t)
  names(A)[1] <- 'Step'
  return(A)
}



## Evaluation
set.seed(2019)
pl.dt <- gather(paths.CIR(), 'Path', 'Value', 2:41)
set.seed(2019)
pl.dt.refl <- gather(paths.CIR(negs = 'reflect'), 'Path', 'Value', 2:41)

ggplot(pl.dt,aes(Step,Value,group=Path)) +
  geom_line(aes(colour=Path))+
  theme(legend.title = element_blank())+
  theme(legend.position = "none")+
  ggtitle('CIR paths - truncated to zero in case of r<0')
```
In the CIR model the diffusion coefficient, $\sqrt{\alpha r_t}$ avoids the possibility of negative interest rates for all parameter values.

As we use Euler discretization we use in this case truncation to avoid negative values.
\newpage
```{r CIR_reflect}
ggplot(pl.dt.refl,aes(Step,Value,group=Path))+
  geom_line(aes(colour=Path))+
  theme(legend.title = element_blank())+
  theme(legend.position = "none")+
  ggtitle('CIR paths - reflected around zero in case of r<0') 

```

Here we use Euler discretization and reflection to avoid negative values.


\newpage
## Exercise 3

We price a Look-back call option for which we know that the maturity $T = 2$ years, the current stock price is $S_0 = 10$, the stock's volatility $\sigma = 0.2$ and assume that $r=0.01$.

We use monthly monitoring, i.e. $\Delta = 1/12$ and Monte Carlo simulation with $10000$ scenarios.

To do so we use naive Monte Carlo for a look-back call option and implement the following Algorithm (we have to generate the whole paths here as our option is path-dependent):
```{r Lookback_Option}
Call.naive.mc <- function(S0, sigma, r, delta, T_years, n) {
  m <- T_years/delta
  S <- matrix(rep(0, (m+1)*n), nrow=n)
  S[,1] <- rep(S0, n)
  SM <- payoff <- c(rep(0,n))
  Z <- matrix(rep(0, m*n), nrow=n)

  for (i in 1:n) {
    for (j in 1:m) {
      Z[i,j] <- rnorm(1, mean = 0, sd = 1)
      S[i,j+1] <- S[i,j]*exp((r-sigma^2/2)*delta +sigma*sqrt(delta)*Z[i,j])
      }
    SM[i] <- max(S[i,])
    payoff[i] <- exp(-r*T_years)*(SM[i]-S[i,m])
  }
  
  price <- mean(payoff)
  se <- sd(payoff)/sqrt(n)
  z.score <- qnorm(1-0.05/2, mean = 0, sd = 1)
  low.b <- price - z.score*se
  up.b <- price + z.score*se
  width <- up.b - low.b
  return(c(MC.naive=price, s.e.=se, Lower=low.b, Upper=up.b, ci.width = width))
}


# inputs
S0 <- 10
sigma <- 0.2
r <- 0.01 # assumed
delta <- 1/12
T_years <- 2
n <- 10000

# Evaluation
set.seed(2019)
Call.naive.mc(S0, sigma, r, delta, T_years, n)
```
\newpage
## Exercise 4

### Part a)

Suppose we have the setting given in page 116 of the slide set. We solve this exercise by using Euler discretization and come up with a Monte Carlo estimate and the corresponding errors.

Setting: Call option pricing under Heston model 

We have call option writen on a stock which is assumed to follow Heston model with parameters $T = 1$, $S_0 = K = 100$, $r = 0.05$, $V_0 = 0.04$, $\alpha = 1.2$ (speed of mean-reversion), $\bar{V} = 0.04$ (long-run volatility level), $\zeta = 0.3$ (diffusion coeff of volatility) and $\rho = -0.5$.

$\rightarrow$ Fourier inversion methods can be used which would yield the price $10.3009$.


Now we price this option by using Monte Carlo with an Euler scheme:
\begin{itemize}
  \item We take $n = 10, n = 50 n = 100, n = 500$ and $n = 1000$ as number of (time) steps and $m = 20000$ as the number of paths.
  \item Additionally we compute the mean absolute error for each case.
\end{itemize}

```{r Heston Call}
# Heston Call pricing (MC - Euler)
Call.Heston.mc.E<-function(S0,K,r,T_years,V0,Vbar,alpha,zeta,rho,m,n,Call.Fourier){
  delta <- T_years/n
  S <- V <- matrix(rep(0, (n+1)*m), nrow=m)
  S[,1] <- rep(S0, m)
  V[,1] <- rep(V0, m)
  payoff <- error <- c(rep(0,m))
  Z <- Z1 <- Zs <- matrix(rep(0, m*n), nrow=m)
  
  for (i in 1:m) {
    Z[i,] <- rnorm(n, mean = 0, sd = 1)
    Z1[i,] <- rnorm(n, mean = 0, sd = 1)
    Zs[i,] <- rho*Z[i,] + sqrt(1-rho^2)*Z1[i,]
    for (j in 1:n) {
      V[i,j+1] <- max(0, V[i,j]+alpha*delta*T_years*(Vbar-V[i,j])+
                         zeta*sqrt(V[i,j]*delta*T_years)*Z[i,j])
      S[i,j+1]<-max(0,S[i,j]*((1+r*delta*T_years)+sqrt(V[i,j]*delta*T_years)*Zs[i,j]))
      }
    payoff[i] <- exp(-r*T_years)*max(S[i,n]-K, 0)
    error[i] <- payoff[i]-Call.Fourier
  }

  a.m.error <- abs(mean(error))
  Price <- mean(payoff)  
  se <- sd(payoff)/sqrt(n)
  z.score <- qnorm(1-0.05/2, mean = 0, sd = 1)
  low.b <- Price - z.score*se
  up.b <- Price + z.score*se
  width <- up.b - low.b
  return(c(MC.Euler=Price, s.e.=se, Lower=low.b, Upper=up.b, 
           ci.width = width, abs.mean.error = a.m.error))
}


# inputs
S0 <- K <- 100 
r <- 0.05
T_years <- 1
V0 <- 0.04
Vbar <- 0.04
alpha <- 1.2
zeta <- 0.3
rho <- -0.5
m <- 20000 #1e06
Call.Fourier <- 10.3009 # Fourier inversion


# Evaluation
set.seed(10)
C.Hes.E.n10 <- Call.Heston.mc.E(S0, K, r, T_years, V0, Vbar, alpha, zeta, rho, m, 
                                Call.Fourier, n = 10); C.Hes.E.n10
set.seed(10)
C.Hes.E.n50 <- Call.Heston.mc.E(S0, K, r, T_years, V0, Vbar, alpha, zeta, rho, m, 
                                Call.Fourier, n = 50); C.Hes.E.n50
set.seed(10)
C.Hes.E.n100 <- Call.Heston.mc.E(S0, K, r, T_years, V0, Vbar, alpha, zeta, rho, m, 
                                 Call.Fourier, n = 100); C.Hes.E.n100
set.seed(10)
C.Hes.E.n500 <- Call.Heston.mc.E(S0, K, r, T_years, V0, Vbar, alpha, zeta, rho, m, 
                                 Call.Fourier, n = 500); C.Hes.E.n500
set.seed(10)
C.Hes.E.n1000 <- Call.Heston.mc.E(S0, K, r, T_years, V0, Vbar, alpha, zeta, rho, m, 
                                  Call.Fourier, n = 1000); C.Hes.E.n1000

abs.mean.errors.E <- c(C.Hes.E.n10[6], C.Hes.E.n50[6], C.Hes.E.n100[6], 
                       C.Hes.E.n500[6], C.Hes.E.n1000[6])
n <- c(10, 50, 100, 500, 1000)
plot(n, abs.mean.errors.E, type = "o"
     ,xlab= "# of iterations n=10,50,100,500,1000",
     main = "Weak error crit. (Heston Call MC-Euler pricing)")
```


\newpage
### Partb)

Now we repeat this exercise where we use the Milstein scheme to price the same call option. We also compute the errors corresponding to different $n$ values and compare them with the ones we have obtained above in a) for the Euler discretization.

We apply the Milstein scheme to the variance process only, as this usually turns out to be the cause of potential problems (See e.g. \url{https://dms.umontreal.ca/~bedard/Heston.pdf} p.11)

```{r Heston_w_Millstein}
# Heston Call pricing (MC - Milstein)
Call.Heston.mc.M<-function(S0,K,r,T_years,V0,Vbar,alpha,zeta,rho,m,n,Call.Fourier){
  delta <- T_years/n
  S <- V <- matrix(rep(0, (n+1)*m), nrow=m)
  S[,1] <- rep(S0, m)
  V[,1] <- rep(V0, m)
  payoff <- error <- c(rep(0,m))
  Z <- Z1 <- Zs <- matrix(rep(0, m*n), nrow=m)
  
  for (i in 1:m) {
    Z[i,] <- rnorm(n, mean = 0, sd = 1)
    Z1[i,] <- rnorm(n, mean = 0, sd = 1)
    Zs[i,] <- rho*Z[i,] + sqrt(1-rho^2)*Z1[i,]
    for (j in 1:n) {
      V[i,j+1] <- max(0, V[i,j]+alpha*delta*T_years*(Vbar-V[i,j])+  
                        # Milstein scheme applied to variance process only
                        zeta*sqrt(V[i,j]*delta*T_years)*Z[i,j]+     
                        # as it usually turns out to be the cause of potential problem
                        0.25*zeta^2*delta*T_years*((Z[i,j])^2-1))
      S[i,j+1] <- max(0, S[i,j]*((1+r*delta*T_years)
                                 +sqrt(V[i,j]*delta*T_years)*Zs[i,j]))
      }
    payoff[i] <- exp(-r*T_years)*max(S[i,n]-K, 0)
    error[i] <- payoff[i]-Call.Fourier
  }
  
  a.m.error <- abs(mean(error))
  Price <- mean(payoff)  
  se <- sd(payoff)/sqrt(n)
  z.score <- qnorm(1-0.05/2, mean = 0, sd = 1)
  low.b <- Price - z.score*se
  up.b <- Price + z.score*se
  width <- up.b - low.b
  return(c(MC.Milstein=Price, s.e.=se, Lower=low.b, Upper=up.b, 
           ci.width = width, abs.mean.error = a.m.error))
}


# inputs
S0 <- K <- 100
r <- 0.05
T_years <- 1
V0 <- 0.04
Vbar <- 0.04
alpha <- 1.2
zeta <- 0.3
rho <- -0.5
m <- 20000 #1e06
Call.Fourier <- 10.3009 # Fourier inversion


# Evaluation
set.seed(10)
C.Hes.M.n10 <- Call.Heston.mc.M(S0, K, r, T_years, V0, Vbar, alpha, zeta, rho, m, 
                                Call.Fourier, n = 10); C.Hes.M.n10
set.seed(10)
C.Hes.M.n50 <- Call.Heston.mc.M(S0, K, r, T_years, V0, Vbar, alpha, zeta, rho, m, 
                                Call.Fourier, n = 50); C.Hes.M.n50
set.seed(10)
C.Hes.M.n100 <- Call.Heston.mc.M(S0, K, r, T_years, V0, Vbar, alpha, zeta, rho, m, 
                                 Call.Fourier, n = 100); C.Hes.M.n100
set.seed(10)
C.Hes.M.n500 <- Call.Heston.mc.M(S0, K, r, T_years, V0, Vbar, alpha, zeta, rho, m, 
                                 Call.Fourier, n = 500); C.Hes.M.n500
set.seed(10)
C.Hes.M.n1000 <- Call.Heston.mc.M(S0, K, r, T_years, V0, Vbar, alpha, zeta, rho, m, 
                                  Call.Fourier, n = 1000); C.Hes.M.n1000

abs.mean.errors.M <- c(C.Hes.M.n10[6], C.Hes.M.n50[6], C.Hes.M.n100[6], 
                       C.Hes.M.n500[6], C.Hes.M.n1000[6])
n <- c(10, 50, 100, 500, 1000)
plot(n, abs.mean.errors.M, type = "o"
     ,xlab= "# of iterations n=10,50,100,500,1000"
     ,main = "Weak error crit. (Heston Call MC-Milstein pricing)")
```

\newpage
### Comparison of Milstein and Euler approximation:

We check the weak order of convergence to get a feeling for the behavior of the error. We see a general decrease in the mean absolute error as the number of time steps increases. The various conditions (on both the option payoff and the SDE) that are required to guarantee a given order of convergence of the schemes is not satisfied, i.e. sometimes Euler will perform better and sometimes Milstein (as we can see in the graph, however the difference is quite small). Overall, the outcome would depend highly on the generated paths.
```{r}
plot(n, abs.mean.errors.M, type = "o", col = "blue"
     ,xlab= "# of iterations n=10,50,100,500,1000"
     ,main = "Comparison: Weak error crit. (Heston Call MC pr.)")
lines(n, abs.mean.errors.E, type = "o", col = "red" )
legend("topright", legend=c("Milstein", "Euler"), 
       col=c("blue", "red"), lty=1, cex=0.8)
```



\newpage
## Exercise 5

### Part a)

Suppose we have the setting given in page 129 (T=2, S_0=100, $\sigma$=0.2, r=0.01, K=100, $\Delta$=1/12, n=10000) of the slide set. We price this Asian option by using control variates method where we choose the standard European call option as the control variate $(Y1 = e^{-rT} (S_{T}-K)^{+})$. Moreover, we provide the corresponding $95\%$ confidence intervals and also compute the price by using only the naive MC for comparison.

We know that $m=T/\Delta$. Furthermore, we chose to take $k=10,000$ for the pilot simulations and $n=100,000$ for the main simulation (which have to be carried out independently):
```{r Asian Option}
## Naive MC 
Call.mc <- function(S0, K, vol, T_years, delta, r, n) {
  m <- T_years / delta
  S1 <- matrix(rep(0, (m+1)*n), nrow=n)
  S1[,1] <- rep(S0, n)
  payoff.A1 <- c(rep(0,n))
  A1 <- c()
  Z4 <- matrix(rep(0, n*m), nrow=n)
  for (i in 1:n) {
    Z4[i,] <- rnorm(m, mean = 0, sd = 1)
    for (j in 1:m) {
      S1[i,j+1] <- S1[i,j]*exp((r-vol^2/2)*delta + vol*sqrt(delta)*Z4[i,j])
    }
    A1[i] <- mean(S1[i,])
    payoff.A1[i] <- exp(-r*T_years)*max(A1[i]-K, 0)
  }

  Price <- mean(payoff.A1) # MC-estimate
  se <- sd(payoff.A1)/sqrt(n)
  z.score <- qnorm(1-0.05/2, mean = 0, sd = 1)
  low.b <- Price - z.score*se
  up.b <- Price + z.score*se
  width <- up.b - low.b
  return(c(MC.est=Price, s.e.=se, Lower=low.b, Upper=up.b, ci.width = width))
}


## MC with control variates
Call.mc.cv <- function(S0, K, vol, T_years, delta, r, k, n) {
  ## Pilot simulation
  m <- T_years / delta
  S <- matrix(rep(0, (m+1)*k), nrow=k)
  S[,1] <- rep(S0, k)
  payoff.A <- c(rep(0,k))
  payoff.E.C <- c()
  A <- c()
  Z1 <- matrix(rep(0, k*m), nrow=k)
  for (i in 1:k) {
    Z1[i,] <- rnorm(m, mean = 0, sd = 1)
    for (j in 1:m) {
      S[i,j+1] <- S[i,j]*exp((r-vol^2/2)*delta 
                             +vol*sqrt(delta)*Z1[i,j]) # Simulate Si values (GBM)
    }
    A[i] <- mean(S[i,])
    payoff.A[i] <- exp(-r*T_years)*pmax(A[i]-K, 0) # Arithmetic-average Asian Call
    payoff.E.C[i] <- exp(-r*T_years)*pmax(S[i,m]-K, 0)# Eur. Call payoff: control var.
  }
  a <- -cov(payoff.A, payoff.E.C)/var(payoff.E.C) # estimate of c*
  corr.X.Y <- cor(payoff.E.C, payoff.A)
  
  ## Main simulation (indep.)
  S2 <- matrix(rep(0, (m+1)*n), nrow=n)
  S2[,1] <- rep(S0, n)
  payoff.A2 <- c(rep(0,n))
  payoff.E.C2 <- c() 
  A2 <- c()
  Z2 <- matrix(rep(0, n*m), nrow=n)
  for (i in 1:n) {
    Z2[i,] <- rnorm(m, mean = 0, sd = 1)
    for (j in 1:m) {
      S2[i,j+1] <- S2[i,j]*exp((r-vol^2/2)*delta + vol*sqrt(delta)*Z2[i,j])
    }
    A2[i] <- mean(S2[i,])
    payoff.A2[i] <- exp(-r*T_years)*max(A2[i]-K, 0)
    payoff.E.C2[i] <- exp(-r*T_years)*pmax(S2[i,m]-K, 0) 
  }

  payoff_cv <- payoff.A2 + a*(payoff.E.C2 - mean(payoff.E.C2)) # controlled estimator
  Price <- mean(payoff_cv) # MC-CV-estimate
  se <- sd(payoff_cv)/sqrt(n)
  z.score <- qnorm(1-0.05/2, mean = 0, sd = 1)
  low.b <- Price - z.score*se
  up.b <- Price + z.score*se
  width <- up.b - low.b
  return(
  c(MC.cv=Price,s.e.=se,Lower=low.b,Upper=up.b,ci.width=width,corr.X.Y=corr.X.Y)
  )
}

## Input
delta <- (1/12)
T_years <- 2
S0 <- 100
vol <- 0.2
r <- 0.01
K <- 100
n <- 100000
k <- 10000

## Evaluations & Comparison
set.seed(10)
naive.mc <- Call.mc(S0, K, vol, T_years, delta, r, n); naive.mc # naive MC
# MC CV with European Call
set.seed(10)
naive.mc.cv.E <- Call.mc.cv(S0, K, vol, T_years, delta, r, k, n); naive.mc.cv.E
```

\newpage
### Part b)

Now we take the geometric-average Asian option as control variate $(Y2 = e^{-rT} max(A_{dg}-K, 0)^{+})$ and compute the price of the arithmetic-average Asian option from before. Again we provide the corresponding $95\%$ confidence intervals and compare this with the result we have obtained above: 
```{r Geom_Asian_Control}
# define geom mean function
gm_mean <- function(x, na.rm=TRUE){
  exp(sum(log(x[x > 0]), na.rm=na.rm) / length(x))
}

## MC with control variates
Call.mc.cv2 <- function(S0, K, vol, T_years, delta, r, k, n) {
  ## Pilot simulation
  m <- T_years / delta
  S1 <- matrix(rep(0, (m+1)*k), nrow=k)
  S1[,1] <- rep(S0, k)
  payoff.G.A <- c(rep(0,k))
  payoff.A.A <- c(rep(0,k))
  A1 <- A2 <- c()
  Z1 <- matrix(rep(0, k*m), nrow=k)
  for (i in 1:k) {
    Z1[i,] <- rnorm(m, mean = 0, sd = 1)
    for (j in 1:m) {
     S1[i,j+1]<-S1[i,j]*exp((r-vol^2/2)*delta + vol*sqrt(delta)*Z1[i,j])
     #Simulate Si's (GBM by drawing from a standard normal dist.)
     }
    A1[i] <- gm_mean(S1[i,])
    payoff.G.A[i]<-exp(-r*T_years)*pmax(A1[i]-K, 0)#Geom-avg Asian Call:payoffs is CV
    A2[i] <- mean(S1[i,])
    payoff.A.A[i]<-exp(-r*T_years)*pmax(A2[i]-K, 0)#Arithmetic-average Asian Call
  }
  a <- -cov(payoff.A.A, payoff.G.A)/var(payoff.G.A) # estimate of c*
  corr.X.Y <- cor(payoff.G.A, payoff.A.A)
  
  # Main simulation (indep.)
  S2 <- matrix(rep(0, (m+1)*n), nrow=n)
  S2[,1] <- rep(S0, n)
  payoff.G.A2 <- c(rep(0,n))
  payoff.A.A2 <- c(rep(0,n))
  A3 <- A4 <- c()
  Z2 <- matrix(rep(0, n*m), nrow=n)
  for (i in 1:n) {
    Z2[i,] <- rnorm(m, mean = 0, sd = 1)
    for (j in 1:m) {
      S2[i,j+1] <- S2[i,j]*exp((r-vol^2/2)*delta + vol*sqrt(delta)*Z2[i,j]) 
    }
    A3[i] <- gm_mean(S2[i,])
    payoff.G.A2[i] <- exp(-r*T_years)*pmax(A3[i]-K, 0) 
    A4[i] <- mean(S2[i,])
    payoff.A.A2[i] <- exp(-r*T_years)*max(A4[i]-K, 0)
  }
  
  payoff_cv<-payoff.A.A2+a*(payoff.G.A2-mean(payoff.G.A2))#controlled estimator
  Price <- mean(payoff_cv) # MC-CV-estimate
  se <- sd(payoff_cv)/sqrt(n)
  z.score <- qnorm(1-0.05/2, mean = 0, sd = 1)
  low.b <- Price - z.score*se
  up.b <- Price + z.score*se
  width <- up.b - low.b
  return(
  c(MC.cv=Price,s.e.=se,Lower=low.b,Upper=up.b,ci.width=width,corr.X.Y=corr.X.Y)
  )
}

## Input
delta <- (1/12)
T_years <- 2
S0 <- 100
vol <- 0.2
r <- 0.01
K <- 100
n <- 100000
k <- 10000

## Evaluations & Comparison
set.seed(10)
naive.mc <- Call.mc(S0, K, vol, T_years, delta, r, n); naive.mc # naive MC
# MC CV with European Call
set.seed(10)
naive.mc.cv.E <- Call.mc.cv(S0, K, vol, T_years, delta, r, k, n); naive.mc.cv.E 
# MC CV with Geometric-average Asian Call
set.seed(10)
naive.mc.cv.GA <- Call.mc.cv2(S0, K, vol, T_years, delta, r, k, n); naive.mc.cv.GA 
```
### Comparison:

The effectiveness of a control variate, as measured by the variance reduction ratio is determined by the strength of the correlation between the quantity of interest $X$ and the control $Y$. Therefore, the higher the correlation the better the control variant (ceteris paribus).

Theoretically, when using the standard European Call as control variate, we capture the nonlinearity in the option payoff, but we do not capture the essential features of the payoff, as we only concentrate on the final value of the stock.

When using the geometric-average Asian option as control variate we capture both, the essential features of the option payoff, as well as the nonlinearity in the option payoff.

This also coincides with our empirical outcomes (compare the two correlations and widths of the confidence intervals).

\newpage
## Exercise 6:

### Part a)
 
```{r Barrier_Option}
# Barrier Option Naive Monte Carlo
Euler.Barrier.MC<-function(S0=60,K=50,b=40,r=0.01,sigma=0.4,Tcap=1/4,m=65,n=10000){
#65 is the num. of days in the current quarter
# iterations for naive monte carlo - plain vanilla
Ps<-rep(NA,n)
S<-as.data.frame(matrix(rep(c(S0,rep(NA,m)),n),ncol=n))
names(S)<-paste0('Path.',c(1:n))
exting<-0 # records the number of knock-outs
for(i in 1:n){
  for(j in 1:(m)){
  S[j+1,i]<-S[j,i]*exp((r-sigma^2/2)*Tcap/m+sigma*sqrt(Tcap/m)*rnorm(1,0,1))
  }
  I_S<-1*(min(S[,i])>b*exp(-0.5826*sigma*sqrt(Tcap/m)))
  exting<-exting+(1-I_S)
  Ps[i]<-exp(-r*Tcap)*(I_S*max(K-S[m+1,i],0))
}
S<-cbind(c(0:m),S)
names(S)[1]<-'Day'
P_est<-mean(Ps)
return(list( P_do=P_est
            ,MC_SE=(sqrt(sum((Ps-P_est)^2)/(n-1)))/sqrt(n)
            ,Knock.Outs=exting
            ,Paths.Frame=S)
       )
}

tic()
n<-5000
Put<-Euler.Barrier.MC()
toc()

# Monte Carlo Estmate of Option Price (in $)
Put$P_do
# Estimator's Standard Error
Put$MC_SE
```

```{r, echo=FALSE}
# Repeat with only 200 paths for a plot
Put.200<-Euler.Barrier.MC(n = 200)
s.pl<-gather(Put.200$Paths.Frame,'Path','Stock_Price',2:201)
ggplot(s.pl,aes(Day,Stock_Price,group=Path))+
geom_rect(aes(xmin=0, xmax=65, ymin=40,ymax=50,group=Path)
          ,color="transparent", fill="cadetblue1", alpha=0.005)+
geom_rect(aes(xmin=0, xmax=65, ymin=30,ymax=40,group=Path)
          ,color="transparent", fill="orange", alpha=0.005)+
geom_line(aes(colour=Path))+
#geom_text(c('Option in the money','KO'),mapping = aes(x=c(1,1),y=c(42,32)))+
theme(legend.title = element_blank())+
theme(legend.position = "none")+
ggtitle('Stock paths')+
annotate(geom="text", x=9, y=45, label="Option in the money",color="darkslategray4")+
annotate(geom="text", x=9, y=35, label="Option knocked-out",color="darkorange4")
```

```{r, eval=FALSE}
Put.200<-Euler.Barrier.MC(n = 200)
s.pl<-gather(Put.200$Paths.Frame,'Path','Stock_Price',2:201)
ggplot(s.pl,aes(Day,Stock_Price,group=Path))+
geom_line(aes(colour=Path))
```
### Part b)

Importance sampling is a useful variance reduction technique for the cases of barrier option pricing, where the event of crossing the set knock-out/knock-in price is highly unlikely. Consider an up-and-in option as an example. In such case, if the barrier level is set high above the initial level of an underlying, naive Monte Carlo will result in most of the simulations wasted, as the resulting pay-off will simply be zero.

In similar instances, it is possible to increase efficiency by introducing some drift $c$ towards the barrier price level and then correcting for this change by multiplication of our estimator by the corresponding likelihood ratio. By assuming the stock price follows Geometric Brownian Motion, the price increments will be log-normally distributed:

$$ logS_{t}-logS_{t-\Delta} = Z_t \sim N\left(\mu,\sigma^2\Delta\right) $$

With $\mu = \left(r-\frac{\sigma^2}{2}\right)\Delta$.

In our step-wise generation in the manner : $S_{j+1}=S{j}\cdot e^{(\mu+\sigma*\sqrt{\Delta}*Z_j)}$ this would amount to setting $\tilde{\mu}=\mu-c$ (with $c>0$ for down-options and $c<0$ for up-options) and to multiplication of the Monte Carlo Estimator $e^{-rT}(I(\mathbf{S})(\pm K\mp S_M)^+)$  \footnote{with $+K-S_M$ for put option and $-K+S_M$ for the call option} by the ratio of two joint density functions:
$$ \hat{\Theta}_{IS}=\frac{f(z_1,..,z_M)}{g(z_1,..,z_M)}\mathbb{E}_g\left[I(\mathbf{S})(\pm K\mp S_M)^+)\right] $$

